BERT Model Compression and Weight Sharing

### 2019-05-13 paper/paper.tex
Delete large comment blocks and TeX code cleanup
First half of paper proofread

\begin{abstract}
% \raggedright
  We apply model compression techniques to fine-tuned BERT on multiclass
  classification problems.  Knowledge distillation and DeepTwist-based compression techniques give
  us moderate compression benefits for a commensurate decrease in model
  performance. We achieve our most compelling results from employing the DeepTwist
  framework to share weights across BERT models fine-tuned for different tasks. Our work demonstrates most weights need not change in the fine-tuning
  process.
%   \item This document describes the expected style, structure, and rough
%   proportions for your final project write-up.
%   \item While you are free to break from this structure, consider it a strong
%   prior for our expectations of the final report.
%   \item Length is a hard constraint. You are only allowed max \textbf{8 pages}
%   in this format. While you can include supplementary material, it will not be
%   factored into the grading process. It is your responsibility to convey the
%   main contributions of the work in the length given.
  % \end{itemize}

@@

% Example Structure:
% \begin{itemize}
% \item What is the problem of interest and what (high-level) are the current best methods for solving it?
% \item How do you plan to improve/understand/modify this or related methods?
% \item Preview your research process, list the contributions you made, and summarize your experimental findings.
% \end{itemize}

@@

Previous work has had a lot of success in computer vision tasks.
\citet{han2015deep} was able to combine pruning, quantization and Huffman
coding to reduce the size of AlexNet by 35$\times$ and the size of VGG-16 by
49$\times$ without loss of accuracy. Recently, similar compression techniques
were applied by \citet{cheong2014transformer} on language models, Transformers
in specific, and reduced the size by a factor of 5.85 while retaining 98.43\%
of the performance. To the best of our knowledge, compressing language models
still remains an under-studied field.

In this paper, we demonstrate that the DeepTwist framework
\citep{lee2018deeptwist} is applicable to compressing fine-tuned BERT
\citep{devlin2018bert} on multiclass classification problems in addition to
computer vision tasks, though the drop in performance is less negligible.
Knowledge distillation also achieves the compression objective with more than
desired decrease in predictive accuracy.

We used two of the GLUE datasets SST-2 and QQP. We ran most of the experiments
on SST-2, a binary sentiment classification task on single-sentence movie
reviews. We validated the results on QQP, a binary classification task for
whether two questions asked on Quora are semantically equivalent.

The most important contribution of this paper is that we proposed a new
weight-sharing method called ``Difference Pruning'' which imposes constraint on
the sparcity of weight difference with the pretrained BERT model. While
uncompressed fine-tuned BERT model achieves 92.9\% validation accuracy on
SST-2, the Difference Pruning model that keeps 99\% of the pretrained BERT
weights achieves 91.3\% on the validation set. This enables multiple language
modeling tasks to share the weights of pretrained BERT and each incurs minimal
extra storage space requirements.

@@

% Example Structure:
% \begin{itemize}
% \item What information does a non-expert need to know about the problem domain?
% \item What data exists for this problem?
% \item What are the challenges/opportunities inherent to the data? (High
% dimensional, sparse, missing data, noise, structure, discrete/continuous,
% etc?)
% \end{itemize}

DNNs have become extremely large in size with growing complexities. A good
example is BERT \citep{devlin2018bert}, which represents a growing
literature in natural language processing in favor of pretraining 
\citep{peters2018deep,radford2018improving,radford2019language}, which
uses large models to learn from a large corpus of text in an unsupervised
manner. The goal of such pretraining is creating contextualized
representation of text in $\R^n$ which preserve certain structure of
language. For different tasks, the researcher then \emph{fine-tunes} the
pretrained model to the task, which may reach good performance since the
pretrained representations are well-behaved representations of language
strucutre. A downside of this framework is that each fine-tuned task
requires a unique set of weights, which, when coupled with the raw size of
these models, are cumbersome objects especially in a low hardware resource
environment.

The smaller \texttt{BERT-base-uncased} model has 110M parameters and takes
438 MB disk space to store. This
makes it very hard to deploy those models on mobile devices, which in most
cases have limited storage and RAM and low computational resources. As a
result, we are not able to store multiple models for different tasks on mobile
devices. Allocating hundreds of megabytes to to those models is also not
realistic for mobile apps such as Gmail or Facebook. And due to increasing
privacy concerns, users prefer models generating predictions based on their
input to be running locally instead of sending information to the cloud. Last
but not least, evaluating enormous models takes too much time on mobile
devices and can drain batteries significantly.

@@

% Example Structure:
% \begin{itemize}
% \item What 3-5 papers have been published in this space?
% \item How do these  differ from your approach?
% \item What data or methodologies do each of these works use?
% \item How do you plan to compare to these methods?
% \end{itemize}

@@

\citet{han2015deep} applies compression techniques to computer vision,
where convolutional neural networks are popular and are often
overparameterized \citep{yang2019fsnet},
where the network is gradually \emph{pruned} and retrained in the first
stage, where small
weights are set
to zero and gradient connections severed. In the second stage, the weights
that survived pruning are clustered, and each cluster behaves like a single
weight, with identical gradients. In the last stage, the model is Huffman
encoded. Numerical experiments by \citet{han2015deep} show that the
compression does not seem to influence performance. Moreover, the benefits
of various compression techniques appear to stack without much loss in
compression benefit. Moreover, convolutional network sizes are dominated by
the convolution operations in
lower layers \citep{denton2014exploiting}. \citet{denton2014exploiting}
uses matrix factorization techniques on the matrices in convolutional
layers, which has low effective rank.

@@

@@

In this paper, we limit our purview to text classification tasks. Training
neural-network-based systems for this task typically consists of a form of
gradient descent based on a differentiable loss function which takes in the
prediction for a given example and the correct label. That is, if we have
training data $(\bm{x}_i, y_i)$ and model $f$, we would aim to minimize,
\[ \ell(f(\bm{x_i}), y_i)\]
where usually the loss function is taken to be Cross-Entropy loss. When we
already have a model $f$ that performs well, we can leverage this training
paradigm through a modification called \emph{knowledge distillation}. Given
the large model that performs well, termed the \emph{teacher network}, we
train a smaller network $g$, the \emph{student network}, by minimizing,
\[ \ell(g(\bm{x_i}), f(\bm{x_i})).\]
Here, we can no longer use Cross-Entropy loss because $f(\bm{x}_i)$ is a
probability distribution rather than a specific label. However, we can take
$\ell$ to be either mean-square error (MSE) between the probabilities in
log-space, or Kullback-Leibler (KL) loss. We avoid MSE error on the
probabilities themselves because of disappearing gradients when the predicted
probabilities are very close to 0 or 1.

@@

An additional benefit from knowledge distillation arises from BERT's
pre-training. Because BERT is pretrained on large amounts of unstructured
data, it contains a powerful language model which likely would be impossible
to learn from the specific task's data-set. To an extent, relevant features
from this language model are what make BERT so powerful once it's fine-tuned.
Training the student network on BERT's output thus allows some of this
``outside knowledge'' to percolate toward the student network.

To provide a concrete example of how this might work, suppose BERT captures
from its pre-training that ``bad'', ``horrible,'' and ``terrible'' are
approximate synonyms. In a sentiment task, these three words appear in three
different sentences, two of which are negative, and one of which is positive.
A fine-tuned BERT may predict a lower likelihood of being positive to the
positive sentence than it would otherwise even if the word of interest appears
nowhere else.

@@

To this end, we follow the framework put forward by \citet{lee2018deeptwist},
\emph{DeepTwist}, whose training loop is characterized by \cref{algo:deeptwist}.
Training of the model closely resembles SGD with one important modification:
every $S_d$ batches, the model weights are transformed by a \emph{distortion}
function.

\begin{algorithm}
  \begin{algorithmic}
    \STATE{$n \leftarrow$ number of batches per epoch}
    \STATE{$\theta \leftarrow$ initialization of model parameters}
    \STATE{$k \leftarrow$ number of epochs to train}
    \STATE{$S_d \leftarrow$ distortion step}
    \FOR{$i=0$ to $k \cdot n$}
      \STATE{batch $\leftarrow$ batches[$i \pmod n$]}
      \STATE{$\theta \leftarrow \theta - \eta \nabla \ell(\theta,
      \text{batch})$}
      \COMMENT{gradient step}
      \IF{$i \equiv 0 \pmod{S_d}$}
        \STATE{$\theta \leftarrow \distortion(\theta)$}
      \ENDIF
    \ENDFOR
    \STATE{$\theta \leftarrow \distortion(\theta)$}
  \end{algorithmic}
  \caption{DeepTwist Training Loop}
  \label{algo:deeptwist}
\end{algorithm}

We have significant flexibility in picking a distortion function, but in order
to be useful, it must provide compression benefits. That is, the image of the
distortion function should have a parameterization that takes less space in
memory than then range of all possible $\theta$s. For example, if $\theta \in
\mathbb{R}^{m,n}$, and the distortion produces a matrix of rank 1, then any
distorted matrix could be uniquely identified by $\theta' \in \mathbb{R}^{n+m}$
through a matrix factorization.

It is important to note that while $\theta$ will be compressible immediately
after a distortion step, there is no guarantee that it will remain so after some
gradient steps. In fact, almost surely it will not be. Thus, we also enforce
that the process ends on a distortion step.

We first set up some notation. Let $F_\btheta = G_\bpsi \circ \bert_\bphi$ be
a transformed model where $\btheta = (\bphi, \bpsi)$ be parameters of the
subparts of the model. $\bert_\bphi$ is a BERT model where the pretrained
weights are $\bphi_0 = (\bphi_0^i)_{i=1}^n$, and $G_\bpsi$ is additional
layers for fine-tuning BERT, parameterized by $\bpsi$. We decompose the
subparts of the model into layers: \begin{align*}
\bert_\bphi &= f_{n}^{\bphi_n} \circ \cdots \circ f_1^{\bphi_1} \\
G_\bpsi &= g_{m}^{\bpsi_m} \circ \cdots \circ g_1^{\bpsi_1}.
\end{align*}

Given training data $(\bm x_i, y_i)_{i=1}^N$, where $\bm x_i$ are one-hot
encoded tokens, we train by solving the standard maximum likelihood problem
where \[
\min_{\btheta} \sum_{i=1}^N \ell\pr{y_i, F_\btheta(\bm x_i)}
\]
for some loss function specific to the task. For instance, $\ell$ is
cross-entropy loss for a (multiclass) classification problem, and $\ell$ is
mean-squared-error for a regression problem, which correspond to
log-likelihoods in Categorical and Gaussian models, respectively. The
innovation of DeepTwist-based training is perturbing the usual gradient-based
optimization by repeatedly distorting the current weights to achieve
compression objectives.

Specifically, the DeepTwist framework presented in \Cref{sec:training} applies
a distortion step for fixed intervals during training, where \[
\btheta \gets \distortion(\btheta). 
\]
In this section, we discuss a few methods for parameterizing $\distortion$.
Generally, we may imagine a few special cases of distortion, we call a
particular distortion \emph{distortion by layer} if it is applied per layer,
i.e. \begin{align*}
\distortion'(\btheta) &= \\
(&\distortion(\bphi_1),\ldots,\distortion(\bphi_n),\\
&\distortion(\bpsi_1),\ldots, \distortion(\bpsi_m)).
\end{align*}
We call a particular distortion \emph{pretrained-only} if \[
\distortion'(\btheta) = \distortion(\bphi).
\]

@@

% \begin{itemize}
% \item What were the results comparing previous work / baseline systems / your
% systems on the main task?
% \item What were the secondary results comparing the variants of your system?
% \item This section should be fact based and relatively dry. What happened,
% what was significant?
% \end{itemize}

% \begin{table*}
%   \centering
%   \missingfigure{}
%   \caption{This is usually a table. Tables with numbers are generally easier
%   to read than graphs, so prefer when possible.}
%   \label{fig:mainres}
% \end{table*}

% \begin{table}[tbh]
%         \caption{Results for DeepTwist}
%         \label{tab:deeptwist_results}
%         \centering
%         \vspace{1em}
%         \begin{tabular}{lccccc}
% \toprule
% {} &               type &  param &  val loss &   val acc \\
% {} & 
% \hypertarget{tabcol:1}{(1)} & \hypertarget{tabcol:2}{(2)} & \hypertarget{tabcol:3}{(3)} & \hypertarget{tabcol:4}{(4)} \\ \midrule
% 0  &  diff prune global &   $90$ &   $0.211$ &   $0.912$ \\
% 1  &              prune &   $20$ &   $0.291$ &   $0.886$ \\
% 2  &              prune &   $80$ &   $0.532$ &   $0.767$ \\
% 3  &         diff prune &   $80$ &   $0.222$ &   $0.916$ \\
% 4  &  diff prune global &   $99$ &       --- &       --- \\
% 5  &  diff prune global &   $90$ &   $0.213$ &   $0.916$ \\
% 6  &         diff prune &   $40$ &   $0.203$ &   $0.909$ \\
% 7  &                rank reduction &   $50$ &   $0.490$ &   $0.810$ \\
% 8  &                rank reduction &  $200$ &   $0.459$ &   $0.845$ \\
% 9  &              prune &   $40$ &   $0.424$ &   $0.845$ \\
% 10 &              prune &   $60$ &   $0.425$ &   $0.815$ \\
% 11 &  diff prune global &   $99$ &   $0.230$ &   $0.901$ \\
% 12 &  diff prune global &   $99$ &   $0.222$ &   $0.905$ \\
% 13 &         diff prune &   $99$ &   $0.222$ &   $0.913$ \\
% 14 &         diff prune &   $95$ &   $0.221$ &   $0.916$ \\
% 15 &         diff prune &   $90$ &   $0.216$ &   $0.916$ \\
% 16 &         diff prune &   $60$ &   $0.209$ &   $0.920$ \\
% 17 &                rank reduction &  $100$ &   $0.515$ &   $0.823$ \\
% 18 &                rank reduction &  $150$ &   $0.456$ &   $0.828$ \\
% 19 &         diff prune &   $80$ &   $2.533$ &   $0.491$ \\
% 20 &         diff prune &   $90$ &       --- &       --- \\
% 21 &  diff prune global &    $0$ &   $0.738$ &   $0.509$ \\
% 22 &         diff prune &    $0$ &   $0.692$ &   $0.522$ \\
% 23 &         diff prune &    $0$ &   $1.854$ &   $0.491$ \\
% \bottomrule
% \end{tabular}

% \end{table}

@@

% \begin{table}[tbh]
%         \caption{Results for Knowledge Distillation}
%         \label{tab:kd_results}
%         \centering
%         \vspace{1em}
%         \begin{tabular}{lccccccc}
% \toprule
% {} &  framework &  type &  val loss &  ß val acc \\
% {} & \hypertarget{tabcol:1}{(1)} & \hypertarget{tabcol:2}{(2)} & \hypertarget{tabcol:3}{(3)} & \hypertarget{tabcol:4}{(4)} \\
% \midrule
% 0 &        nokd &   CNN &   $0.906$ &   $0.799$ \\
% 1 &        nokd &   CNN &   $0.448$ &   $0.784$ \\
% 2 &        nokd &   CNN &   $0.856$ &   $0.812$ \\
% 3 &          kd &  LSTM &   $0.738$ &   $0.831$ \\
% 4 &          kd &   CNN &   $0.423$ &   $0.834$ \\
% \bottomrule
% \end{tabular}
% \label{tbl:kdresults}
% \end{table}

@@

% Things to say about DeepTwist
% \begin{itemize}
%   \item Stratified prune better
%   \item Prune rate mad robust
%   \item Diff gud
%   \item ``Improved performance''
%   \item QQP takes long, but looks okay
%   \item Didn't try only pruning the $\bert$ part
% \end{itemize}
% 
% Things to say about KD
% \begin{itemize}
%   \item it bad
%   \item doesn't work
%   \item LSTM weird/CNN incredibly noisy
%   \item \TODO{compare}
% \end{itemize}

@@

% \begin{itemize}
% \item What conclusions can you draw from the results section?
% \item Is there further analysis you can do into the results of the system? Here
% is a good place to include visualizations, graphs, qualitative analysis of your
% results.
% 
% \item  What questions remain open? What did you think might work, but did not?
% \end{itemize}

% \begin{figure}
%   \centering
%   \missingfigure{}
%   \missingfigure{}
%   \missingfigure{}
%   \caption{Visualizations of the internals of the system.}
% \end{figure}

@@

% TODO: thank Rush + 287

% \textbf{Do not} include acknowledgements in the initial version of
% the paper submitted for blind review.

% If a paper is accepted, the final camera-ready version can (and
% probably should) include acknowledgements. In this case, please
% place such acknowledgements in an unnumbered section at the
% end of the paper. Typically, this will include thanks to reviewers
% who gave useful comments, to colleagues who contributed to the ideas,
% and to funding agencies and corporate sponsors that provided financial
% support.

>>>>

\begin{abstract}
% \raggedright
  We apply model compression techniques to fine-tuned BERT on multiclass
  classification problems.  Knowledge distillation and DeepTwist-based
  compression techniques give us moderate compression benefits for a
  commensurate decrease in model performance. We achieve our most compelling
  results from employing the DeepTwist framework to share weights across
  BERT-based
  models fine-tuned for different tasks. Our work demonstrates most weights need
  not change in the fine-tuning process.%   \item This document describes the expected style, structure, and rough
%   proportions for your final project write-up.
%   \item While you are free to break from this structure, consider it a strong
%   prior for our expectations of the final report.
%   \item Length is a hard constraint. You are only allowed max \textbf{8 pages}
%   in this format. While you can include supplementary material, it will not be
%   factored into the grading process. It is your responsibility to convey the
%   main contributions of the work in the length given.
  % \end{itemize}

@@

@@

Previous work has had a lot of success in computer vision tasks.
\citet{han2015deep} was able to combine pruning, quantization and Huffman
coding to reduce the size of AlexNet by 35$\times$ and the size of VGG-16 by
49$\times$ without loss of accuracy. Recently, similar compression techniques
were applied by \citet{cheong2014transformer} on language models, Transformers
especially, and reduced the size by a factor of 5.85 while retaining
98.43\%
of the performance. To the best of our knowledge, compressing language models
still remains an under-studied field.

In this paper, we demonstrate that the DeepTwist framework
\citep{lee2018deeptwist} is applicable to compressing fine-tuned BERT
\citep{devlin2018bert} on multiclass classification problems in addition to
computer vision tasks, though the drop in performance is less negligible.
Knowledge distillation also achieves the compression objective with more
sacrifices in  predictive accuracy.

We use two of the GLUE datasets SST-2 and QQP. We conduct most of the
experiments
on SST-2, a binary sentiment classification task on single-sentence movie
reviews. We validate the results on QQP, a binary classification task for
whether two questions asked on Quora are semantically equivalent.

Our most important contribution is a new weight-sharing method called
\emph{difference pruning}, which imposes constraint on the sparsity of weight
difference with the pretrained BERT model. While uncompressed fine-tuned BERT
model achieves 92.9\% validation accuracy on SST-2, the difference pruning model
that keeps 99\% of the pretrained BERT weights achieves 91.3\% on the validation
set, which is on par with the fully fine-tuned model. This enables multiple
language modeling tasks to share the weights of pretrained BERT, each incurring
minimal extra storage space requirements.

@@

DNNs have become extremely large in size with growing complexities. A good
example is BERT \citep{devlin2018bert}, which epitomizes a growing
literature in natural language processing in favor of pretraining 
\citep{peters2018deep,radford2018improving,radford2019language},
using large models to learn from a large corpus of text in an unsupervised
manner. The goal of such pretraining is creating contextualized
representation of text in $\R^n$ which preserve certain structure of
language. For different tasks, the researcher then \emph{fine-tunes} the
pretrained model to the task, which may reach good performance since the
pretrained representations are well-behaved representations of language
strucutre. A downside of this framework is that each fine-tuned task
requires a unique set of weights, which, when coupled with the raw size of
these models, are cumbersome objects especially in a low hardware resource
environment.

The smaller \texttt{BERT-base-uncased} model has 110M parameters and takes
438 MB of disk space to store. This
makes it very hard to deploy these models on mobile devices, which in most
cases have limited storage, RAM, and computational resources. As a
result, we are not able to store multiple models for different tasks on mobile
devices. Allocating hundreds of megabytes to to those models is also not
realistic for mobile apps such as Gmail or Facebook. And due to increasing
privacy concerns, users prefer models generating predictions based on their
input to be running locally instead of sending information to the cloud. Last
but not least, evaluating enormous models takes too much time on mobile
devices and can drain batteries significantly.

@@

@@

\citet{han2015deep} applies compression techniques to computer vision, where
convolutional neural networks are popular and are often overparameterized
\citep{yang2019fsnet}, where the network is gradually \emph{pruned} and
retrained in the first stage, by setting small weights to zero and severing
gradient connections. In the second stage, the weights that survived pruning are
clustered, and each cluster behaves like a single weight, with identical
gradients. In the last stage, the model is Huffman encoded. Numerical
experiments by \citet{han2015deep} show that the compression does not seem to
influence performance. Moreover, the benefits of various compression techniques
appear to stack without much loss in compression benefit. Additionally,
convolutional network sizes are dominated by the convolution operations in lower
layers \citep{denton2014exploiting}. \citet{denton2014exploiting} uses matrix
factorization techniques on the matrices in convolutional layers, which has low
effective rank.

@@

We first set up some notation. Let $F_\btheta = G_\bpsi \circ \bert_\bphi$ be
a fine-tuned model where $\btheta = (\bphi, \bpsi)$ are parameters of the
subparts of the model. $\bert_\bphi$ is a BERT model where the pretrained
weights are $\bphi_0 = (\bphi_0^i)_{i=1}^n$, and $G_\bpsi$ is additional
layers for fine-tuning BERT, parameterized by $\bpsi$. We decompose the
subparts of the model into layers: \begin{align*}
\bert_\bphi &= f_{n}^{\bphi_n} \circ \cdots \circ f_1^{\bphi_1} \\
G_\bpsi &= g_{m}^{\bpsi_m} \circ \cdots \circ g_1^{\bpsi_1}.
\end{align*}

Given training data $(\bm x_i, y_i)_{i=1}^N$, where $\bm x_i$ are one-hot
encoded tokens, we train by solving the standard maximum likelihood problem
where the objective is \[
\min_{\btheta} \sum_{i=1}^N \ell\pr{F_\btheta(\bm x_i), y_i}
\]
for some loss function specific to the task. For instance, $\ell$ is
cross-entropy loss for a (multiclass) classification problem, and $\ell$ is
mean-squared-error for a regression problem, which correspond to
log-likelihoods in Categorical and Gaussian models, respectively.

@@

In this paper, we limit our purview to text classification tasks. Training
neural-network-based systems for this task typically consists of
gradient-descent-based optimizers applied to a differentiable loss
function, which takes in the prediction for a given example and the correct label. That is, if we have
training data $(\bm{x}_i, y_i)$ and model $f$, we would aim to minimize,
\[ \ell(f(\bm{x_i}), y_i)\]
where usually the loss function is taken to be cross-entropy loss. When we
already have a model $f$ that performs well, we can leverage this training
paradigm through a modification called \emph{knowledge distillation}. Given
the large model that performs well, termed the \emph{teacher network}, we
train a smaller network $g$, the \emph{student network}, by minimizing,
\[ \ell(g(\bm{x_i}), f(\bm{x_i})).\]
We can take
$\ell$ to be either mean-square error (MSE) between the probabilities in
log-space, or Kullback-Leibler (KL) loss. We avoid MSE error on the
probabilities themselves because of disappearing gradients when the predicted
probabilities are very close to 0 or 1.

@@

An additional benefit from knowledge distillation arises from BERT's
pretraining. Because BERT is pretrained on large amounts of unstructured
data, it contains a powerful language model which likely would be impossible
to learn from the specific task's dataset. To an extent, relevant features
from this language model are what make BERT so powerful once it's fine-tuned.
Training the student network on BERT's output thus allows some of this
``outside knowledge'' to percolate toward the student network.

To provide a concrete example of how this might work, suppose BERT captures
from its pretraining that ``bad'', ``horrible,'' and ``terrible'' are
approximate synonyms. In a sentiment task, these three words appear in three
different sentences, two of which are negative, and one of which is positive.
A fine-tuned BERT may predict a lower likelihood of being positive to the
positive sentence than it would otherwise even if the word of interest appears
nowhere else.

@@

To this end, we follow the framework put forward by \citet{lee2018deeptwist},
\emph{DeepTwist}, whose training loop is characterized by \cref{algo:deeptwist}.
Training of the model closely resembles normal gradient training with one
important modification: every $S_d$ batches, the model weights are transformed
by a \emph{distortion} function.

\begin{algorithm}
  \begin{algorithmic}
    \STATE{$n \leftarrow$ number of batches per epoch}
    \STATE{$\btheta \leftarrow$ initialization of model parameters}
    \STATE{$k \leftarrow$ number of epochs to train}
    \STATE{$S_d \leftarrow$ distortion step}
    \FOR{$i=0$ to $k \cdot n$}
      \STATE{batch $\leftarrow$ batches[$i \pmod n$]}
      \STATE{$\btheta \leftarrow \btheta - \eta \nabla \ell(\btheta,
      \text{batch})$}
      \COMMENT{gradient step}
      \IF{$i \equiv 0 \pmod{S_d}$}
        \STATE{$\btheta \leftarrow \distortion(\btheta)$}
      \ENDIF
    \ENDFOR
    \STATE{$\btheta \leftarrow \distortion(\btheta)$}
  \end{algorithmic}
  \caption{DeepTwist Training Loop}
  \label{algo:deeptwist}
\end{algorithm}

We have significant flexibility in picking a distortion function, but in order
to be useful, it must provide compression benefits. That is, the image of the
distortion function should have a parameterization that takes less space in
memory than then range of all possible $\btheta$s. For example, if $\btheta \in
\mathbb{R}^{m,n}$, and the distortion produces a matrix of rank 1, then any
distorted matrix could be uniquely identified by $\btheta' \in \mathbb{R}^{n+m}$
through a matrix factorization.

It is important to note that while $\btheta$ will be compressible immediately
after a distortion step, there is no guarantee that it will remain so after some
gradient steps. In fact, almost surely it will not be. Thus, we also enforce
that the process ends on a distortion step.

\section{Distortions}
\label{sec:distortions}
The innovation of DeepTwist-based training is perturbing the usual
gradient-based optimization by repeatedly distorting the current weights to
achieve compression objectives. Specifically, the DeepTwist framework presented
in \Cref{sec:training} applies a distortion step for fixed intervals during
training, where \[
\btheta \gets \distortion(\btheta). 
\]
In this section, we discuss a few methods for parameterizing $\distortion$.
Generally, we may imagine a few special cases of distortion, we call a
particular distortion \emph{distortion by layer} if it is applied per layer,
i.e. \begin{align*}
\distortion'(\btheta) &= \\
(&\distortion(\bphi_1),\ldots,\distortion(\bphi_n),\\
&\distortion(\bpsi_1),\ldots, \distortion(\bpsi_m)).
\end{align*}
We call a particular distortion \emph{pretrained-only} if \[
\distortion'(\btheta) = \distortion(\bphi).
\]

@@

@@

@@

@@

@@
